---
title: "Leaflet2"
output: html_document
---

```{r setup, include=FALSE}
#Install these packages

require(rgdal)
require(leaflet)
require(leaflet.extras)
require(dplyr)
require(readxl)
require(stringr)
require(rgeos)
require(sp)
```

#Map 1 focuses on having multiple choices for basemap layers | The data consists of a feature class for all adult homes in Washington state. 

#Reading data 
```{r}
Homes <- read.csv('Datafiles/DSHS_Adult_Family_Homes.csv')
#head(Homes)
```

#add layerControls()
```{r}


leaflet(data = Homes) %>%
  addProviderTiles(providers$Esri.NatGeoWorldMap, options = providerTileOptions(noWrap = TRUE), group = "NGC") %>%
  addProviderTiles(providers$Stamen.Toner, options = providerTileOptions(noWrap = TRUE), group = "Stamen") %>%
  addMarkers(~Lat, ~Long) %>%
  addLayersControl(
    baseGroups = c("NGC", "Stamen"),
overlayGroups = c("Homes", "WASeniors"),
options = layersControlOptions(collapsed = FALSE)
)

```

#Map 2 is a mapping of a line layer on to the basemap | Feature layer is a layer of railroads in LongIsland, NY

# Lines
```{r, lines}
rail <- readOGR("./Datafiles/New_York_State_Railroad_Lines.shp")

leaflet(data = rail) %>%
  addProviderTiles("Stamen.Watercolor", options = providerTileOptions(noWrap = TRUE)) %>%
  addPolylines(color = "brown")


```


#Map 3 deals with polygons | We utilize the crime data from hw1 and merge that with the US states Shape file to create a new dataset. We then map different crimes and state populations on to the map

#Create new table for polygon 
```{r}
states <- readOGR("./Datafiles/cb_2018_us_state_20m.shp")
crime <- read.csv("./Datafiles/data.csv")
#head(crime)
#head(states)
```

#Merge Tables to form new shape file with crime data
```{r}

StateCrime <- states[states$NAME %in% crime$State,]

StateCrime@data <- merge(StateCrime@data, crime, sort = FALSE, by.x = "NAME", by.y = "State")
#head(StateCrime)
```


#Create element bins 
```{r}
bins1 <- c(0,10000000, 25000000, Inf)

pal1 <- colorBin(
  palette = "Purples",
  domain = StateCrime$Pop, 
  bins = bins1)

bins2 <- c(0,1000, 2500, Inf)

pal2 <- colorBin(
  palette = "Reds",
  domain = StateCrime$firearmDeaths, 
  bins = bins2)

bins3 <- c(0,5 ,10, Inf)

pal3 <- colorBin(
  palette = "Blues",
  domain = StateCrime$homicideRate2017,
  bins = bins3)

```


#mapping the feature layer on to the map

```{r}


leaflet(data = StateCrime) %>%

  addProviderTiles("Stamen.Toner", options = providerTileOptions(noWrap = TRUE)) %>%
  addPolygons(color = ~pal1(`Pop`), popup = ~paste0("<b>", `Pop`, ":</b> ", round(`Pop`, 2), " years"), group = "Population", weight = 1, smoothFactor = 0.5,
    opacity = 1.0, fillOpacity = 0.8) %>%
  addLegend(position = "bottomright", pal = pal1, values = StateCrime$`Pop`, title = "State Population", group = "Population")%>%
  addPolygons(color = ~pal2(`firearmDeaths`), popup = ~paste0("<b>", `firearmDeaths`, ":</b> ", round(`firearmDeaths`, 2), " years"), group = "FirearmDeaths", weight = 1, smoothFactor = 0.5,
    opacity = 1.0, fillOpacity = 0.8) %>%
  addLegend(position = "bottomright", pal = pal2, values = StateCrime$`firearmDeaths`, title = "State Deaths by Firearms", group = "FirearmDeaths")%>%
  addPolygons(color = ~pal3(`homicideRate2017`), popup = ~paste0("<b>", `homicideRate2017`, ":</b> ", round(`homicideRate2017`, 2), " years"), group = "Homicide", weight = 1, smoothFactor = 0.5,
    opacity = 1.0, fillOpacity = 0.8) %>%
  addLegend(position = "bottomright", pal = pal3, values = StateCrime$`homicideRate2017`, title = "State Homicide Rate", group = "Homicide")%>%
  addLayersControl(
      overlayGroups  = c("Population", "FirearmDeaths", "Homicide"),
options = layersControlOptions(collapsed = FALSE)
)
  
```

